---
title: "MSDS 6306 - Case Study 02"
author: "Huy Hoang Nguyen"
date: "12/05/2019"
output:
  html_document: default
pdf_document: default
---




```{r libraries, echo=F, results='hide', message=F, warning=F}
library(tidyverse) #The "tidyverse" collects some of the most versatile R packages: ggplot2, dplyr, tidyr, readr, purrr, and tibble. The packages work in harmony to clean, process, model, and visualize data.
library(skimr) #for data summary - so sweet and I like a lot this library
library(mice) #package provides a nice function md.pattern() to get a better understanding of the pattern of missing data
library(VIM) #more helpful visual representation can be obtained using the VIM package for agrr
library(naniar) #https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html (for gg_mis_var) (Missing values)
library(mlbench) #collection of artificial and real-world machine learning benchmark problems, including, e.g., several data sets from the UCI repository. (also has BostonHousing)
library(caret)
library(mlr)
library(tidyverse)
library(ggthemes)
library(gplots)
library(randomForest)
library(corrplot)
library(kableExtra)
library(plotly)
library(GGally) #for ggpairs
library(Boruta) #for Automated EDA later
library(readxl) #read excel
library(e1071) #Naive Bayes

```


```{r data loading, echo=F, results='hide', message=F, warning=F}


rawdata0 <- read.csv("CaseStudy2-data.csv")
head(rawdata0)
view(rawdata0) #There are 870 entries, 36 total columns
length(rawdata0) #[1] 36
skim(rawdata0) #so sweet 0- for data summary

rawdata <- rawdata0
```


## VII. EDA5 - Multiple Linear Regression and Validation for Salary 


Here we will use our EDA4 to build model for Monthly Income.


### 1. Build a Model   


I will build a multiple linear regression on Monthly Income related to 14 important variables in the last part.



```{r}

set.seed(100)
data3 <- select(rawdata, "MonthlyIncome", "Age", "Attrition", "BusinessTravel", "Department", "Education", "JobLevel", "JobRole", "NumCompaniesWorked", "TotalWorkingYears", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager")
splitPerc1 = 0.8
trainIndices1 = sample(1:dim(data3)[1],round(splitPerc1 * dim(data3)[1]))
train1 = data3[trainIndices1,]
test1 = data3[-trainIndices1,]
lm1 = lm(MonthlyIncome ~ ., data = train1)
summary(lm1)
pred1 = predict(lm1, test1)
RMSE = sqrt(mean((test1$MonthlyIncome - pred1)^2))
RMSE


```


Then RMSE is around $1031.816.   



### 2. Validation Requirement for Salary


Here I will use the file CaseStudy2CompSet No Salary.xlsx and create new dataframe case2nosalary1  with the same variables as the dataframe data3 in the last step (Build a Model) (without MonthlyIncome variable).


```{r}
case2nosalary <- read_excel("CaseStudy2CompSet No Salary.xlsx")
case2nosalary1 <- select(case2nosalary, "Age", "Attrition", "BusinessTravel", "Department", "Education", "JobLevel", "JobRole", "NumCompaniesWorked", "TotalWorkingYears", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager")
```


Now I will use  the model built in the last step to predicting the salary for the file CaseStudy2CompSet No Salary.xlsx. 



```{r}

lm2 = lm(MonthlyIncome ~ ., data = data3)
pred2 = predict(lm2, case2nosalary1)


MonthlyIncome = pred2
MonthlyIncome = as.data.frame(MonthlyIncome)

case2nosalary2 = cbind(case2nosalary, MonthlyIncome)


write.csv(case2nosalary2, "Case2PredictionsNguyenSalary.csv", row.names = FALSE)

case2nosalary3 <- read.csv("Case2PredictionsNguyenSalary.csv", header = T)

head(case2nosalary3)

```

Later, I will upload the preditecd file Case2PredictionsNguyenSalary.csv for Salary into github. 



