---
title: "MSDS 6306 - Case Study 02"
author: "Huy Hoang Nguyen"
date: "12/05/2019"
output:
  html_document: default
pdf_document: default
---


```{r libraries, echo=F, results='hide', message=F, warning=F}
library(tidyverse) #The "tidyverse" collects some of the most versatile R packages: ggplot2, dplyr, tidyr, readr, purrr, and tibble. The packages work in harmony to clean, process, model, and visualize data.
library(skimr) #for data summary - so sweet and I like a lot this library
library(mice) #package provides a nice function md.pattern() to get a better understanding of the pattern of missing data
library(VIM) #more helpful visual representation can be obtained using the VIM package for agrr
library(naniar) #https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html (for gg_mis_var) (Missing values)
library(mlbench) #collection of artificial and real-world machine learning benchmark problems, including, e.g., several data sets from the UCI repository. (also has BostonHousing)
library(caret)
library(mlr)
library(tidyverse)
library(ggthemes)
library(gplots)
library(randomForest)
library(corrplot)
library(kableExtra)
library(plotly)
library(GGally) #for ggpairs

```


```{r data loading, echo=F, results='hide', message=F, warning=F}
rawdata <- read.csv("https://github.com/hnguye01/hnguye01.github.io/raw/master/DS6306/Data/CaseStudy2-data.csv")
head(rawdata)
view(rawdata) #There are 870 entries, 36 total columns
length(rawdata) #[1] 36
skim(rawdata) #so sweet 0- for data summary
```

```{r dropping columns, echo=F, results='hide', message=F, warning=F}
drop_columns <- which(apply(rawdata, 2, function(x) (length(unique(x)) == 1)))

cols <- names(drop_columns)
rawdata <- rawdata[,-drop_columns]

#Actually, we can drop manually by another code as  rawdata <- select(rawdata, -c("Over18","EmployeeCount", "StandardHours")) . We will get the same results finally.

skim(rawdata)
```

```{r dropping more columns, echo=F, results='hide', message=F, warning=F}
rawdata <- select(rawdata, -c("ID","EmployeeNumber"))
skim(rawdata)
```

```{r data0, echo=F, results='hide', message=F, warning=F}
factorcolumns <- c("JobInvolvement", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction", "WorkLifeBalance")

rawdata[,factorcolumns] <- lapply(rawdata[,factorcolumns], as.factor)
data0 <- rawdata #data0 - dataset that I use for the analysis
skim(data0)

```

```{r data2 , echo=F, results='hide', message=F, warning=F}
data2 <- select(data0, -c("HourlyRate","DailyRate", "MonthlyRate","PerformanceRating"))
```


## IV. EDA2 - T-test Analysis

### 1. t-test Analysis on Attrition variable:


```{r Age}
t.Age <- t.test(Age~Attrition, data=data2)
t.Age
t.MonthlyIncome <-  t.test(MonthlyIncome~Attrition, data=data2)
t.MonthlyIncome
t.Education <-  t.test(Education~Attrition, data=data2)
t.Education
t.DistanceFromHome <-  t.test(DistanceFromHome~Attrition, data=data2)
t.DistanceFromHome
t.EnvironmentSatisfaction <-  t.test(EnvironmentSatisfaction~Attrition, data=data2)
t.EnvironmentSatisfaction
t.JobLevel <-  t.test(JobLevel~Attrition, data=data2)
t.JobLevel
t.NumCompaniesWorked  <-  t.test(NumCompaniesWorked ~Attrition, data=data2)
t.NumCompaniesWorked 
t.PercentSalaryHike  <-  t.test(PercentSalaryHike ~Attrition, data=data2)
t.PercentSalaryHike
t.StockOptionLevel  <-  t.test(StockOptionLevel ~Attrition, data=data2)
t.StockOptionLevel
t.TotalWorkingYears  <-  t.test(TotalWorkingYears ~Attrition, data=data2)
t.TotalWorkingYears
t.TrainingTimesLastYear  <-  t.test(TrainingTimesLastYear ~Attrition, data=data2)
t.TrainingTimesLastYear
t.YearsAtCompany  <-  t.test(YearsAtCompany ~Attrition, data=data2)
t.YearsAtCompany
t.YearsInCurrentRole  <-  t.test(YearsInCurrentRole ~Attrition, data=data2)
t.YearsInCurrentRole
t.YearsSinceLastPromotion  <-  t.test(YearsSinceLastPromotion ~Attrition, data=data2)
t.YearsSinceLastPromotion
t.YearsWithCurrManager <-  t.test(YearsWithCurrManager ~Attrition, data=data2)
t.YearsWithCurrManager

```



```{r}
testnames <- c("Age","Monthly Income", "Education",
               "Distance From Home", "Environment Satisfaction", 
               "Job Level", "Number of Companies Worked",
               "Percent Salary Hike", "Stock Option Level",
               "Total Working Years", "Training Times Last Year",
               "Years At Company", "Years In Current Role",
               "Years Since Last Promotion", "Years With Current Manager"
               )
testpval <- c(t.Age$p.value,
              t.MonthlyIncome$p.value,
              t.Education$p.value,
              t.DistanceFromHome$p.value,
              t.EnvironmentSatisfaction$p.value,
              t.JobLevel$p.value,
              t.NumCompaniesWorked$p.value,
              t.PercentSalaryHike$p.value,
              t.StockOptionLevel$p.value,
              t.TotalWorkingYears$p.value,
              t.TrainingTimesLastYear$p.value,
              t.YearsAtCompany$p.value,
              t.YearsInCurrentRole$p.value,
              t.YearsSinceLastPromotion$p.value,
              t.YearsWithCurrManager$p.value
              )
ttestout <- cbind.data.frame(testnames,testpval)
ttestout$testpval <- round(ttestout$testpval,10) 
names(ttestout) <- c("Variable","P-Value")
```


The p-values of t-test analyses related to Attrition variable:


```{r}
ttestout %>% kable()
```



- There are two groups for Attrition variable (Yes and No).   
- At $\alpha$-level of significance = 0.05, if p-value<0.05 then we reject the null hypothesis (There are some evidence to suggest that the mean difference in the selected variable for two groups Yes-No of Attrition variable) and if p-value>0.05 then we fail to reject the null hypothesis (There are not enough evidence to suggest that the mean difference in the selected variable for two groups Yes-No of Attrition variable).  
- By  t-test analysis, we will have some ideas to build our models later.  


### 2. t-test Analysis related to Gender variable:


We now will  use t-test to see the difference of Monthly Income in Job role between Male and Female.  

```{r}
summary(data2$JobRole)
table(data2$Gender,data2$JobRole)
```

```{r filter job role}
data2.Healthcare <- data2 %>% filter(JobRole=="Healthcare Representative")
data2.HR <- data2 %>% filter(JobRole=="Human Resources")
data2.Technician <- data2 %>% filter(JobRole=="Laboratory Technician")
data2.Manager <- data2 %>% filter(JobRole=="Manager")
data2.MDirector <- data2 %>% filter(JobRole=="Manufacturing Director")
data2.RDirector <- data2 %>% filter(JobRole=="Research Director")
data2.Scientist <- data2 %>% filter(JobRole=="Research Scientist")
data2.SExecutive <- data2 %>% filter(JobRole=="Sales Executive")
data2.SRepresentative <- data2 %>% filter(JobRole=="Sales Representative")
```

```{r}
t.jHealthcare <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jHealthcare 

t.jHR <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jHR 

t.jTechnician <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jTechnician 

t.jManager <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jManager

t.jMDirector <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jMDirector

t.jRDirector <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jRDirector

t.jScientist <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jScientist

t.jSExecutive <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jSExecutive

t.jSRepresentative <- t.test(MonthlyIncome ~ Gender, data = data2.Healthcare)
t.jSRepresentative

```



```{r}
testnames1 <- c("Healthcare Representative", "Human Resources", "Laboratory Technician",
               "Manager", "Manufacturing Director", "Research Director", "Research Scientist",
               "Sales Executive", "Sales Representative")
testpval1 <- c(t.jHealthcare$p.value,
              t.jHR$p.value,
              t.jTechnician$p.value,
              t.jManager$p.value,
              t.jMDirector$p.value,
              t.jRDirector$p.value,
              t.jScientist$p.value,
              t.jSExecutive$p.value,
              t.jSRepresentative$p.value
              )
ttestout1 <- cbind.data.frame(testnames1,testpval1)
ttestout1$testpval1 <- round(ttestout1$testpval1,10) 
names(ttestout1) <- c("Variable","P-Value")
```



The p-values of t-test analyses related to Monthly Income in Job role between Male and Female:



```{r}
ttestout1 %>% kable()
```



- There are two groups Male and Female for Gender variable.   
- At $\alpha$-level of significance = 0.05 and  p-value>0.05 then we fail to reject the null hypothesis (There are not enough evidence to suggest that the mean difference in Monthly Income related to Job Role for two groups Male-Female of Gender variable).  
- By  t-test analysis, we will have some ideas to build our models later.  
