---
title: "MSDS 6306 - Case Study 02"
author: "Huy Hoang Nguyen"
date: "12/05/2019"
output:
  html_document: default
pdf_document: default
---


```{r libraries, echo=F, results='hide', message=F, warning=F}
library(tidyverse) #The "tidyverse" collects some of the most versatile R packages: ggplot2, dplyr, tidyr, readr, purrr, and tibble. The packages work in harmony to clean, process, model, and visualize data.
library(skimr) #for data summary - so sweet and I like a lot this library
library(mice) #package provides a nice function md.pattern() to get a better understanding of the pattern of missing data
library(VIM) #more helpful visual representation can be obtained using the VIM package for agrr
library(naniar) #https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html (for gg_mis_var) (Missing values)
library(mlbench) #collection of artificial and real-world machine learning benchmark problems, including, e.g., several data sets from the UCI repository. (also has BostonHousing)
library(caret)
library(mlr)
library(tidyverse)
library(ggthemes)
library(gplots)
library(randomForest)
library(corrplot)
library(kableExtra)
library(plotly)
library(GGally) #for ggpairs
library(Boruta)

```


```{r data loading, echo=F, results='hide', message=F, warning=F}


rawdata0 <- read.csv("CaseStudy2-data.csv")
head(rawdata0)
view(rawdata0) #There are 870 entries, 36 total columns
length(rawdata0) #[1] 36
skim(rawdata0) #so sweet 0- for data summary

rawdata <- rawdata0
```

```{r dropping columns, echo=F, results='hide', message=F, warning=F}
drop_columns <- which(apply(rawdata, 2, function(x) (length(unique(x)) == 1)))

cols <- names(drop_columns)
rawdata <- rawdata[,-drop_columns]

#Actually, we can drop manually by another code as  rawdata <- select(rawdata, -c("Over18","EmployeeCount", "StandardHours")) . We will get the same results finally.

skim(rawdata)
```

```{r dropping more columns, echo=F, results='hide', message=F, warning=F}
rawdata <- select(rawdata, -c("ID","EmployeeNumber"))
skim(rawdata)
```

```{r data0, echo=F, results='hide', message=F, warning=F}
factorcolumns <- c("JobInvolvement", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction", "WorkLifeBalance")

rawdata[,factorcolumns] <- lapply(rawdata[,factorcolumns], as.factor)
data0 <- rawdata #data0 - dataset that I use for the analysis
skim(data0)

```

```{r data2 , echo=F, results='hide', message=F, warning=F}
data2 <- select(data0, -c("HourlyRate","DailyRate", "MonthlyRate","PerformanceRating"))




data1 <- data0

factorcolumns1 <- c("Education", "EnvironmentSatisfaction", "JobLevel", "NumCompaniesWorked", "PercentSalaryHike",  "StockOptionLevel", "TotalWorkingYears", "TrainingTimesLastYear", "YearsAtCompany", "YearsInCurrentRole",  "YearsSinceLastPromotion", "YearsWithCurrManager")
data1[,factorcolumns1] <- lapply(data1[,factorcolumns1], as.factor)
```


## VI. EDA4 - Automated EDA

### 1. Analysis on Monthly Income


Now I will use Automated EDA for Feature Selection. I use the library Boruta here to select important variables related to MonthlyIncome variable. 
(http://r-statistics.co/Variable-Selection-and-Importance-With-R.html)  




```{r, echo=T, message=F, warning=F}
boruta_output <- Boruta(MonthlyIncome ~ ., data=data2, doTrace=2)
boruta_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %in% c("Confirmed", "Tentative")])  # collect Confirmed and Tentative variables
print(boruta_signif)  # significant variables
plot(boruta_output, cex.axis=.7, las=2, xlab="", main="Variable Importance")
```

 
Then after running this code, I have 14 important variables as follows: Age,              Attrition, BusinessTravel, Department, Education, JobLevel, JobRole, NumCompaniesWorked, TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager.   


### 2. Analysis on Attrition

Similarly, I will run the following code 


```{r, echo=T, message=F, warning=F}
boruta_output <- Boruta(Attrition ~ ., data=rawdata0, doTrace=2)
boruta_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %in% c("Confirmed", "Tentative")])  # collect Confirmed and Tentative variables
print(boruta_signif)  # significant variables
plot(boruta_output, cex.axis=.7, las=2, xlab="", main="Variable Importance")
```


Then I will use important variable to build a model later: Age, Department,  EnvironmentSatisfaction,  JobInvolvement,  JobLevel, JobRole,  JobSatisfaction, MaritalStatus,  MonthlyIncome, NumCompaniesWorked,  OverTime, StockOptionLevel,  TotalWorkingYears, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole, YearsWithCurrManager.  








