---
title: "MSDS 6306 - Case Study 02"
author: "Huy Hoang Nguyen"
date: "12/05/2019"
output:
  html_document: default
pdf_document: default
---




```{r libraries, echo=F, results='hide', message=F, warning=F}
library(tidyverse) #The "tidyverse" collects some of the most versatile R packages: ggplot2, dplyr, tidyr, readr, purrr, and tibble. The packages work in harmony to clean, process, model, and visualize data.
library(skimr) #for data summary - so sweet and I like a lot this library
library(mice) #package provides a nice function md.pattern() to get a better understanding of the pattern of missing data
library(VIM) #more helpful visual representation can be obtained using the VIM package for agrr
library(naniar) #https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html (for gg_mis_var) (Missing values)
library(mlbench) #collection of artificial and real-world machine learning benchmark problems, including, e.g., several data sets from the UCI repository. (also has BostonHousing)
library(caret)
library(mlr)
library(tidyverse)
library(ggthemes)
library(gplots)
library(randomForest)
library(corrplot)
library(kableExtra)
library(plotly)
library(GGally) #for ggpairs
library(Boruta) #for Automated EDA later
library(readxl) #read excel
library(e1071) #Naive Bayes

```


```{r data loading, echo=F, results='hide', message=F, warning=F}


rawdata0 <- read.csv("CaseStudy2-data.csv")
head(rawdata0)
view(rawdata0) #There are 870 entries, 36 total columns
length(rawdata0) #[1] 36
skim(rawdata0) #so sweet 0- for data summary

rawdata <- rawdata0
```


## VIII. EDA6 - Naive Bayes classifiers and Validation for Attrition


Here I will use our EDA4 to build model for Attrition. I will use these vatiables: Age, Department,  EnvironmentSatisfaction,  JobInvolvement,  JobLevel, JobRole,  JobSatisfaction, MaritalStatus,  MonthlyIncome, NumCompaniesWorked,  OverTime, StockOptionLevel,  TotalWorkingYears, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole, YearsWithCurrManager. 

### 1. Build a Model   


I will use Naive Bayes classifiers to build a model here.


```{r}


data4 <- select(rawdata, "Attrition", "Age", "Department",  "EnvironmentSatisfaction",  "JobInvolvement",  "JobLevel", "JobRole",  "JobSatisfaction", "MaritalStatus",  "MonthlyIncome", "NumCompaniesWorked",  "OverTime", "StockOptionLevel",  "TotalWorkingYears", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsWithCurrManager")


set.seed(10000)
splitPercNB1 = 0.8
trainIndicesNB1 = sample(1:dim(data4)[1],round(splitPercNB1 * dim(data4)[1]))
trainNB1 = data4[trainIndicesNB1,]
testNB1 = data4[-trainIndicesNB1,]

NB1 <- naiveBayes(Attrition~.,data = trainNB1,laplace = -1)
predNB1 = predict(NB1, testNB1)

table(as.factor(testNB1$Attrition),predNB1)

#Confusion Matrix
confusion.Matrix = confusionMatrix(predNB1,as.factor(testNB1$Attrition))
confusion.Matrix
Sensitivity = confusion.Matrix$byClass['Sensitivity']
Specificity = confusion.Matrix$byClass['Specificity']
Accuracy = confusion.Matrix$overall['Accuracy']

Accuracy
Sensitivity
Specificity







```

### 2. Validation Requirement for Attrition  


Here I will use the file CaseStudy2CompSet No Attrition.csv and create new dataframe case2noattrition1  with the same variables as the dataframe data4 in the last step (Build a Model) (without Attrition variable).


```{r}
case2noattrition <- read.csv("CaseStudy2CompSet No Attrition.csv")
case2noattrition1 <- select(case2noattrition, "Age", "Department",  "EnvironmentSatisfaction",  "JobInvolvement",  "JobLevel", "JobRole",  "JobSatisfaction", "MaritalStatus",  "MonthlyIncome", "NumCompaniesWorked",  "OverTime", "StockOptionLevel",  "TotalWorkingYears", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsWithCurrManager")
```


Now I will use  the model built in the last step to predicting Attrition for the file CaseStudy2CompSet No Attrition.csv. 



```{r}


set.seed(10000)

NB2 <- naiveBayes(Attrition~.,data = data4,laplace = -1)
predNB2 = predict(NB2, case2noattrition1)


Attrition = predNB2
Attrition = as.data.frame(Attrition)
case2noattrition2 = cbind(case2noattrition,Attrition)

write.csv(case2noattrition2,"Case2PredictionsNguyenAttrition.csv", row.names = FALSE)

case2noattrition3 <- read.csv("Case2PredictionsNguyenAttrition.csv", header = T)

head(case2noattrition3)
```




Later, I will upload the preditecd file Case2PredictionsNguyenAttrition.csv for Salary into github. 

