---
title: "MSDS 6306 - Case Study 02"
author: "Huy Hoang Nguyen"
date: "12/05/2019"
output:
  html_document: default
pdf_document: default
---


```{r libraries, echo=F, results='hide', message=F, warning=F}
library(tidyverse) #The "tidyverse" collects some of the most versatile R packages: ggplot2, dplyr, tidyr, readr, purrr, and tibble. The packages work in harmony to clean, process, model, and visualize data.
library(skimr) #for data summary - so sweet and I like a lot this library
library(mice) #package provides a nice function md.pattern() to get a better understanding of the pattern of missing data
library(VIM) #more helpful visual representation can be obtained using the VIM package for agrr
library(naniar) #https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html (for gg_mis_var) (Missing values)
library(mlbench) #collection of artificial and real-world machine learning benchmark problems, including, e.g., several data sets from the UCI repository. (also has BostonHousing)
library(caret)
library(mlr)
library(tidyverse)
library(ggthemes)
library(gplots)
library(randomForest)
library(corrplot)
library(kableExtra)
library(plotly)
library(GGally) #for ggpairs

```


```{r data loading, echo=F, results='hide', message=F, warning=F}
rawdata <- read.csv("https://github.com/hnguye01/hnguye01.github.io/raw/master/DS6306/Data/CaseStudy2-data.csv")
head(rawdata)
view(rawdata) #There are 870 entries, 36 total columns
length(rawdata) #[1] 36
skim(rawdata) #so sweet 0- for data summary
```

```{r dropping columns, echo=F, results='hide', message=F, warning=F}
drop_columns <- which(apply(rawdata, 2, function(x) (length(unique(x)) == 1)))

cols <- names(drop_columns)
rawdata <- rawdata[,-drop_columns]

#Actually, we can drop manually by another code as  rawdata <- select(rawdata, -c("Over18","EmployeeCount", "StandardHours")) . We will get the same results finally.

skim(rawdata)
```

```{r dropping more columns, echo=F, results='hide', message=F, warning=F}
rawdata <- select(rawdata, -c("ID","EmployeeNumber"))
skim(rawdata)
```

```{r data0, echo=F, results='hide', message=F, warning=F}
factorcolumns <- c("JobInvolvement", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction", "WorkLifeBalance")

rawdata[,factorcolumns] <- lapply(rawdata[,factorcolumns], as.factor)
data0 <- rawdata #data0 - dataset that I use for the analysis
skim(data0)

```

```{r data2 , echo=F, results='hide', message=F, warning=F}
data2 <- select(data0, -c("HourlyRate","DailyRate", "MonthlyRate","PerformanceRating"))
```


## V. EDA3 - Numeric variables correlation

The following  table will show us the correlations between numeric variables.

```{r}

correlator  <-  function(df){
 df %>%
    keep(is.numeric) %>%
    tidyr::drop_na() %>%
    cor %>%
    corrplot("upper", addCoef.col = "white", number.digits = 2,
             number.cex = 0.5, method="square",
             order="hclust", 
             tl.srt=50, tl.cex = 0.5)
}
correlator(data2)

#Actually, we can remove tidyr::drop_na() %>%  because there is no missing data in this dataset

#We can write the same code directly without using function as follows
#data2 %>% keep(is.numeric) %>% na.omit %>% cor %>% corrplot("upper", addCoef.col = "white", number.digits = 2, number.cex = 0.5, method="square", order="hclust", tl.srt=50, tl.cex = 0.5)

#We can remove na.omit %>% because there is no missing data in this dataset

```

### 1. Analysis on Monthly Income

By this correlation table, we can see  

- MonthlyIncome vs. JobLevel = 0.95  
- MonthIncome vs. TotalWorkingYears = 0.78  
- MonthlyIncome vs. YearsAtCompany = 0.49  
- MonthlyIncome vs. Age = 0.48  
- MonthlyIncome vs. YearsInCurrentRole = 0.36  
- MonthlyIncome vs. YearsWithCurrManager = 0.33  
- MonthlyIncome vs. YearsSinceLastPromotion = 0.32  

Then by some analyses, we can see that Monthly Income have a strong correlation with Job Level. Month Income, Total Working Years and Job Level are correlated strongly. These correlations are logical.


### 2. Analysis on Job Level


- JobLevel vs. MonthlyIncome = 0.95  
- JobLevel vs. TotalWorkingYears = 0.78  
- JobLevel vs. YearsAtCompany = 0.52  
- JobLevel vs. Age = 0.48  
- JobLevel vs. YearsinCurrentRole = 0.39  
- JobLevel vs. YearsWithCurrManager = 0.37  
- JobLevel vs. YearsSinceLastPromotion = 0.33  

Then by some analyses, we can see that Job Level have a strong correlation with Monthly Income. Month Income, Total Working Years and Job Level are correlated strongly. These correlations are logical.  


```{r , echo=F, results='hide', message=F, warning=F}
data2 %>% keep(is.numeric) %>% na.omit %>% cor %>% view()
```









